---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# STA 517 3.0 Programming and Statistical Computing with R

# Generating Random Numbers Using the Inverse Transform Method

Prepared by Dr Thiyanga Talagala

# 1.  Probability distribution functions in R to generate random numbers

\begin{table}[!h]
\begin{tabular}{ll|ll}
\textbf{rbeta} & beta distribution & \textbf{rlnorm} & log-normal distribution \\
\textbf{rbinom} & binomial distribution & \textbf{rmultinom} & multinomial distribution  \\
\textbf{rcauchy} & Cauchy distribution & \textbf{rnbinom} & negative binomial distribution \\
\textbf{rchisq} & chi-squared distribution & \textbf{rnorm}& normal distribution  \\
\textbf{rexp} & exponential distribution &  \textbf{rpois} & Poisson distribution \\
\textbf{rf} & F distribution &  \textbf{rt} & Student's t distribution\\
\textbf{rgamma} & gamma distribution &  \textbf{runif} & uniform distribution\\
\textbf{rgeom} & geometric distribution &  \textbf{rweibull} & Weibull distribution\\
\textbf{rhyper} & hyper-geometric distribution & 
\end{tabular}
\end{table}

There are other methods of generating random numbers from a particular distribution. In this lectorial we will discuss **Inverse Transform Method**.

# 2. Inverse transform method


## Theorem 1: Probability Integral Transformation

> Let $X$ have continuous cdf $F_X(x)$ and define the random variable $Y$ as $Y =  F_X(X)$. Then $Y$ is uniformly distributed on (0, 1), that is, $P(Y \leq y) = y, \text{ 0 < y < 1}.$


Let's try to understand the theorem using an example.

```{r, out.width="100%", fig.asp=0.6, fig.align='center', fig.pos='h', echo=FALSE}
library(ggplot2)
ggplot() + 
  theme_void() + 
  theme(panel.border = element_rect(colour="white", fill=NA, size = 1))

```

\newpage

### Useful results to prove the theorem.

**Result 1:** 

If $F_X$ is strictly increasing, then $F_X^{-1}$ is well defined by 

$$F_X^{-1}(y) = x \Leftrightarrow F_X(x) = y.$$ 

If $F_X$ is constant on some interval,  then $F_X^{-1}$ is not well defined by the above equation. To avoid this problem we define $F_X^{-1}(y)$ for $0 < y < 1$ by

$$F_X^{-1}(y) = inf\{x: F_X(x) \geq y\}.$$


\newpage

**Result 2:** 

If $F_X$ is **strictly** increasing, then it is true that

$$F_X^{-1}(F_X(x)) = x.$$
```{r, out.width="100%", fig.asp=0.4, fig.align='center', fig.pos='h', echo=FALSE}
library(ggplot2)
ggplot() + 
  theme_void() + 
  theme(panel.border = element_rect(colour="white", fill=NA, size = 1))

```


### Proof of Theorem 1:

For $Y = F_X(X)$ we have, for $0 < y < 1,$

\newpage
We can use Theorem 1 to generate random numbers from a particular distribution.

\newpage

# 3. Steps in deriving random numbers using integral transformation method



1. Derive the cumulative distribution function of $f_X(x)$

2. Derive the inverse function $F_X^{-1}(u)$.

3. Write a function to generate random numbers.

    - Generate $u$ from $Uniform(0, 1)$.
    
    - compute $x = F_X^{-1}(u)$.


## Example 1

Write a function to generate $n$ random numbers from the distribution with density $f_X(x) = 3x^2, \text{ 0 < x < 1}.$

**Step 1:** Find the cumulative distribution function of $f_X(x)$,

$$F_X(x) = x^3 \text{ for } 0<x<1$$


\newpage


**Step 2:** Next we need to compute $F_X^{-1}(u)$,

$$F_X^{-1}(u) = u^{\frac{1}{3}}.$$

```{r, out.width="100%", fig.asp=0.9, fig.align='center', fig.pos='h', echo=FALSE}
library(ggplot2)
ggplot() + 
  theme_void() + 
  theme(panel.border = element_rect(colour="white", fill=NA, size = 1))

```

**Step 3:** R function


```{r, comment=NA}
generate_it <- function(n){ 
  # Generate random numbers
  u <- runif(n)
  xgen <- u^(1/3)
  xgen
  
}

set.seed(2020)
generate_it(10)
  
```


**Visualisation of  theoretical distribution**

```{r, comment=NA, fig.height=3, message=FALSE, warning=FALSE}
library(tidyverse)
# Theoretical distribution values
theoretical.df <- tibble(x = seq(0, 1, 0.01), fx = 3*x^2)
ggplot(theoretical.df, aes(x = x, y = fx)) + 
  geom_line(col = "red")
```


**Visualize empirical distribution - counts**

```{r, comment=NA, message=FALSE, fig.height=3}
empirical.df <- data.frame(data.emp = generate_it(1000))  
# Plot empirical distribution - counts
ggplot(empirical.df, aes(x = data.emp))+
  geom_histogram(col = "white", binwidth = 0.05)
  
```

\newpage

**Visualize empirical distribution - density**

```{r, comment=NA, message=FALSE, fig.height=3}
ggplot(empirical.df, aes(x = data.emp, y=..density..)) +
  geom_histogram(col = "white", binwidth = 0.05)
  
```



**Visualize theoretical distribution and empirical distribution together**

```{r, comment=NA, message=FALSE, fig.height=3}
ggplot(empirical.df, aes(x = data.emp, y=..density..)) + 
  geom_histogram(col = "white", binwidth = 0.05) + 
  geom_line(data = theoretical.df, aes(x = x, y = fx), color = 'red') 
```

\newpage

**Function to generate random numbers and visualize theoretical and empirical distributions**

```{r, comment=NA}
generate_it_dist <- function(n){ 
  # Generate random numbers
  u <- runif(n)
  xgen <- u^(1/3)
  xgen
  
  # values for empirical distribution
  empirical.df <- data.frame(xgen=xgen) 
  
  # values for the theoretical distribution
  theoretical.df <- tibble(x = seq(0, 1, 0.01), 
  fx = 3*x^2)
  
  # arrange values and plot into a list
  list(
    xgen,
  ggplot2::ggplot(empirical.df, aes(x=xgen, y=..density..)) + 
    geom_histogram(col="white", binwidth = 0.01) + 
  geom_line(data = theoretical.df, aes(x = x, y = fx), color = 'red') )
}


```

Run the following codes and check the outputs.

```r
# Sample size 10
generate_it_dist(10)

# Sample size 100
n100 <- generate_it_dist(100)
n100
n100[[1]]
n100[[2]]

# Sample size 10000
n10000 <- generate_it_dist(10000)
n10000[[2]]
```

\newpage

**Example 2 ** 


i) Write a function to generate random numbers from the $Exponential(\lambda)$ distribution using the inverse transformation method.


```{r,  comment=NA}
exp_generator <- function(n, lambda){
  u <- runif(n, 0, 1)
  exp.values <- -log(1-u)/lambda
    exp.values
}

```


ii) Generate 1000 random numbers from the $Exponential(2)$ distribution.

```{r, comment=NA, echo=TRUE}
set.seed(111)
exp2 <- exp_generator(1000, 2)

```


```{r, comment=NA}
exp2
```


iii) Graph the density histogram of the sample with the $Exponential(2)$ density superimposed for comparison. 


```{r, comment=NA}
# values for empirical distribution
empirical.df.exp <- data.frame(xgen=exp2) 
# values for the theoretical distribution
theoretical.df <- tibble(x = seq(0.1, 5, 0.01), 
  fx = 2*exp(-2*x))

ggplot(empirical.df.exp, aes(x = xgen, y=..density..)) + 
  geom_histogram(col = "white", binwidth = 0.05) + 
  geom_line(data = theoretical.df, aes(x = x, y = fx), color = "forestgreen") + 
  labs(x="x", y="fx")
```


