---
title: "STA 517 3.0 Programming and Statistical Computing with R"
subtitle: "Functionals - part 1: Overview"
author: "Dr Thiyanga Talagala"
date: "8 November 2020"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: 
      - default
      - default-fonts
      - duke-blue
      - hygge-duke
      - libs/cc-fonts.css
      - libs/figure-captions.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(emo)
```

# Functionals

> A functional is a function that takes a function as an input and returns a vector as output.

> Hadley Wickham, Advanced R

```{r, comment=NA}
statistic <- function(f){
  data <- c(10, 20, 30, 40, 62, 63)
  f(data)
}

```

```{r, comment=NA}
statistic(mean)
statistic(sum)
```


---

# Use of functionals: lapply

lapply: loop over a list and evaluate a function on each element.

```{r, comment=NA}
x <- list( a = 1:8, b = c(2.1, 3.2, 4.2, 5, 6))
x

```

```{r, comment=NA}
#We are passing `mean` is an argument to lapply
lapply(x, mean)
lapply(x, sum)
```


---
# Use of functionals: lapply (cont.)

```{r, comment=NA, message=FALSE}
cv <- function(data){sd(data)/mean(data)}
lapply(x, cv)
```

---
# lapply is a for-loop replacement

```{r, comment=NA}
x <- list( a = 1:8, b = c(2.1, 3.2, 4.2, 5, 6))
x

```

```{r, comment=NA, message=FALSE}
result_x <- list()
result_x
for (i in 1:2){
 result_x[[i]] <- mean(x[[i]])
  
}
result_x
```

---
# Use of functionals: sapply

lapply: loop over a list and evaluate a function on each element.

```{r, comment=NA}
x <- list( a = 1:8, b = c(2.1, 3.2, 4.2, 5, 6))
x

```

```{r, comment=NA}
#We are passing `mean` is an argument to lapply
sapply(x, mean)
sapply(x, sum)
```

Same as `lapply` but the output is a vector.
---

# `map()` function in purrr

```{r, comment=NA}
library(purrr)
```

```{r, comment=NA}
x <- list( a = 1:8, b = c(2.1, 3.2, 4.2, 5, 6))
x

map(x, mean)

```

- The base equivalent to `map()` is `lapply()`.

---
# `map` syntax

> map(YOUR_LIST, YOUR_FUNCTION)

```{r, comment=NA}
map(c(4, 9, 16), sqrt)
```
---

# `map()` 

<!--Before we go on to explore more map variants, let’s take a quick look at how you tend to use multiple purrr functions to solve a moderately realistic problem: fitting a model to each subgroup and extracting a coefficient of the model. For this toy example, I’m going to break the mtcars data set down into groups defined by the number of cylinders, using the base split function. -->

```{r, comment=NA, message=FALSE}
iris_species <- split(iris, iris$Species)
str(iris_species)
```

This creates a list of three data frames.

<!--I think this code is easy to read because each line encapsulates a single step, you can easily distinguish the functional from what it does, and the purrr helpers allow us to very concisely describe what to do in each step.-->


---

# `map()`  (cont.)

```{r, comment=NA, message=FALSE}

head(iris_species)
```

---
# `map()`  (cont.)

Fit a linear model between `Petal.Width ~ Petal.Length` and extract the second coefficient.

```{r, comment=NA, message=FALSE, warning=FALSE}
library(magrittr)
iris_species %>%
  map(~lm(Petal.Width ~ Petal.Length, data = .x)) %>% 
  map(coef) %>%
  map_dbl(2)

```

---


# With the tidyverse

```{r, comment=NA}
iris_species %>%
  map(~lm(Petal.Width ~ Petal.Length, 
          data = .x)) %>% 
  map(coef) %>%
  map_dbl(2)

```


# With base R

```{r, comment=NA}
models <- lapply(iris_species, function(data) lm(Petal.Width ~ Petal.Length,
    data = data))
vapply(models, function(x) coef(x)[[2]], double(1))

```

Or use for loop


---

# With the tidyverse

```{r, comment=NA, message=FALSE, warning=FALSE}
iris_species %>%
  map(~lm(Petal.Width ~ Petal.Length, 
          data = .x)) %>% 
  map(coef) %>%
  map_dbl(2)

```


# For loop

```{r, comment=NA}
intercepts <- double(length(iris_species))
for (i in seq_along(iris_species)) {
  model <- lm(Petal.Width ~ Petal.Length, 
              data = iris_species[[i]])
  intercepts[[i]] <- coef(model)[[2]]
}
intercepts

```

---
# `map` variants: `modify`

```{r, comment=NA}
df <- data.frame(
  x = 1:3,
  y = 6:4
)

map(df, ~ .x * 10)
# If you want to keep the output as a data frame, you can use modify()
modify(df, ~ .x * 10)
```

`modify` returns the same type of output as the input.


---
# `map_df`

```{r, comment=NA}
map_output <- map(mtcars, function(x) length(unique(x)))
head(map_output, 3)
```


```{r, comment=NA}
map_df_output <- map_df(mtcars, function(x) length(unique(x)))
head(map_df_output)
```
---
# `map`/`map_df`/`modify`

```{r, comment=NA}
set.seed(2020)
x <- list(a=rnorm(5), b=rnorm(6))
map(x, mean)
```


```{r, comment=NA}
map_df(x, mean)
```

```{r, comment=NA}
modify(x, mean)
```
---

class: center, middle



Slides available at: https://thiyanga.netlify.app/courses/rmsc2020/contentr/

All rights reserved by [Thiyanga S. Talagala](https://thiyanga.netlify.com/)


Reference: Advanced R, Hadley Wickham
