---
title: "STA 326 2.0 Programming and Data Analysis with R"
subtitle: "R Data Import and Export"
author: "Dr Thiyanga Talagala"
date: "Online distance learning/teaching materials during the COVID-19 outbreak."
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: 
      - default
      - default-fonts
      - duke-blue
      - hygge-duke
      - libs/cc-fonts.css
      - libs/figure-captions.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(emo)
```

# Data import with readr

## R package

`readr`: part of the core tidyverse.

```{r,comment=NA}
library(tidyverse)
```

## `readr` data import functions

- `read_csv`: reads comma-delimited files.

- `read_csv2`: reads semicolon-separated files

- `read_tsv`: reads tab-delimited files


---

# 🛠 Import data from a .csv file


## Syntax


```r
datasetname <- read_csv("include_file_path")
```

When you run `read_csv`, it prints out the names and type of each column.

.full-width[.content-box-yellow[Switch to R]]

  
---

# If the file is saved inside the project folder

.full-width[.content-box-green[Demo: Go to google classroom and watch the video importdatacsv1.mov]]


# If the file is saved outside the project folder

.full-width[.content-box-green[Demo: Go to google classroom and watch the video importdatacsv2.mov]]

---

# 🛠 Importing csv file from a website

## Syntax

```r
datasetname <- read_csv("include url here")
```

## Example

```{r, comment=NA, cache=TRUE}
url <- "https://thiyanga.netlify.app/project/datasets/foodlabel.csv"
foodlabel <- read_csv(url)
```

```{r,comment=NA}
head(foodlabel, 1)
```
---

# `read.csv` and `read_csv`

*  `read.csv` is in base R.

*  `read_csv` is in tidyverse.

*  `read.csv()` performs a similar job to `read_csv()`.

* `read_csv()` works well with other parts of the tidyverse.

*  `read_csv()` is faster than `read.csv()`.

*  `read_csv()` will always read variables containing text as character variable.  In contrast, the base R function `read.csv()` will, by default, convert any character variable to a factor.

<!--This is often not what you want, and can be overridden by passing the option stringsAsFactors = FALSE to read.csv().-->


---

# 🛠 Writing to a File

- We can save tibble (or dataframe) to a csv file, using `write_csv()`.

- `write_csv()` is in the `readr` package.


## Syntax

```r
write_csv(name_of_the_data_set_you_want_to_save, "path_to_write_to")
```

## Example

```r
data(iris)
# This will save inside your project folder
write_csv(iris, "iris.csv") 
# This will save inside the data folder which is inside your project folder
write_csv(iris, "data/iris.csv") 
```

.full-width[.content-box-yellow[Switch to R]]


.full-width[.content-box-green[Demo: Go to google classroom and watch the video exportdatacsv.mov]]


---

# 🛠 Importing Excel .xlsx files

## Syntax

```r
library(readxl)
mydata <- read_xlsx("file_path")
```

.full-width[.content-box-yellow[Switch to R]]


.full-width[.content-box-green[Demo: Go to google classroom and watch the video readxlsx.mov]]

---
# Importing SAS, SPSS and STATA files

## SAS

```r
read_sas("mtcars.sas7bdat")
write_sas(mtcars, "mtcars.sas7bdat")
```
## SPSS

```r
read_sav("mtcars.sav")
write_sav(mtcars, "mtcars.sav")

```
## Stata
```r
read_dta("mtcars.dta")
write_dta(mtcars, "mtcars.dta")
```

---

# Importing other types of data


- `feather`: for sharing with Python and other languages

- `httr`: for web apis

- `jsonlite`: for JSON

- `rvest`: for web scraping

- `xml2`: for XML

.full-width[.content-box-blue[Working with feather, httr, jsonlite, rvest and xml2 is beyond the scope of the course.]]

---

class: center, middle

Slides available at: hellor.netlify.app

All rights reserved by [Thiyanga S. Talagala](https://thiyanga.netlify.com/)



