---
title: "Functionals - Tutorial"
author: "Thiyanga Talagala"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required packages

```{r, comment=NA, message=FALSE, warning=FALSE}
library(tidyverse)
library(purrr)
```

## Dataset

```{r, comment=NA}
library(palmerpenguins)
data(penguins, package = 'palmerpenguins')
head(penguins)
```

Extract numeric columns only


```{r}

penguins.numeric <- penguins %>% select(-c(species, island, sex))
```



## Help

Explore the following command before attempt the question.

```{r, comment=NA}
# Create a blank vector of size 2
output1 <- vector("double", 2)
output1

# Create a blank list of size 3
output2 <- vector("list", 3)
output2

# find number of unique values is a vector
a <- c(1, 1, 2, 3, 4, 5)
n_distinct(a)
```

## Useful map function

`map` and `map_dbl`

## Questions

1. Write code using for loop to compute the mean of every column in `penguins.numeric`.

```{r, comment=NA}
output <- vector("double", ncol(penguins.numeric))# create a blank vector to store the results
names(output) <- names(penguins.numeric)
for (i in names(penguins.numeric)) {
  output[i] <- mean(penguins.numeric[[i]], na.rm = TRUE)
}
output


```

2. Write code that uses one of the map functions to compute the mean of every column in `penguins.numeric`.


```{r, comment=NA}
map_dbl(penguins.numeric, mean, na.rm=TRUE)

```

3. Write for loop to generate 10 random normals for each of
$\mu = 10, 20, 30, 40, 50$.

```{r, comment=NA}
n <- 10
# values of the mean
mu <- c(10, 20, 30, 40, 50)
normals <- vector("list", length(mu))
for (i in 1:length(normals)) {
  normals[[i]] <- rnorm(n, mean = mu[i])
}
normals

```


4. Write code that uses one of the map functions to generate 10 random normals for each of
$\mu = 10, 20, 30, 40, 50$.

```{r, comment=NA}
map(c(10, 20, 30, 40, 50), ~rnorm(n = 10, mean = .))

```


5. Write code that used for loop to compute the number of unique values in each column of the `penguins.numeric` dataset.


```{r, comment=NA, message=FALSE}
penguins.numeric_uniq <- vector("double", ncol(penguins.numeric))
names(penguins.numeric_uniq) <- names(penguins.numeric)
for (i in names(penguins.numeric_uniq)) {
  penguins.numeric_uniq[i] <- n_distinct(penguins.numeric[[i]])
}
penguins.numeric_uniq
```

6. Write code that uses one of the map functions to compute the number of unique values in each column of the `penguins.numeric` dataset.

```{r, comment=NA}
map_dbl(penguins.numeric, n_distinct)

```

*This tutorial is based on R4DS.*

 