---
title: "STA 326 2.0 Programming and Data Analysis with R"
subtitle: "Lesson 5.1: Recap-Functions"
author: "Dr Thiyanga Talagala"
date: "2020-03-10"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: 
      - default
      - default-fonts
      - duke-blue
      - hygge-duke
      - libs/cc-fonts.css
      - libs/figure-captions.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Problem 1

Write a function to prints out standard random normal numbers (use rnorm()) but stops (breaks) if you get a number bigger than 1.


Using next adapt the loop from last exercise so that doesn’t print negative numbers.

---

# Problem 2

Write a function to simulate the flip a coin $n$ times, keeping track of the individual outcomes (1 = heads, 0 = tails) in a vector that
you preallocte. Assume $P(H)=0.5$.

Example output:

```{r, comment=NA, echo=FALSE}
flip_coin <- function(n){
  out <- numeric(n)
  for (i in 1:n){
    coin.out <- rbinom(1, p=0.5, n=1)
    if(coin.out==1){
      out[i] <- "H"
    } else {
      out[i] <- "T"
    }
  }
  out
}

```

```{r, comment=NA}
flip_coin(10)
```


---
```{r, comment=NA}
flip_coin <- function(n){
  out <- numeric(n)
  for (i in 1:n){
    coin.out <- rbinom(1, p=0.5, n=1)
    if(coin.out==1){
      out[i] <- "H"
    } else {
      out[i] <- "T"
    }
  }
  out
}

```

```{r, comment=NA}
flip_coin(10)
```

---

- Modify the function so that it takes another argument, which will be the $P(H)$.

- Run the function 10, 100, 1000 and 10000 times with $P(H)=0.8$ and report the proportion of times that head is selected.

---


# Problem 3

Write a function to calculate the median.

help:

```{r, comment=NA}
5%%2

4%%2
```

---

# Problem 4

Write a function to calculate the correlation coefficient

$$r=\frac{\sum_{i=1}^{n}(x_i-\bar{x})(y_i-\bar{y})}{\sqrt{\sum_{i=1}^n(x_i-\bar{x})^2\sum_{i=1}^n(y_i-\bar{y})^2}}$$

---

class: center, middle

Slides available at: hellor.netlify.com

All rights reserved by [Thiyanga S. Talagala](https://thiyanga.netlify.com/)




