---
title: "Customizing boxplot with ggplot"
author: Thiyanga Talagala
date: '2018-07-05'
slug: blog
categories: [ggplot2]
tags: [R, ggplot2]
---

Boxplot is probably one of the most common type of graphics. This will show how to customize boxplots. 

You will learn,

- how to change the order of labels

- how to change the colours

- how to flip coordinates



Step 1: we create a simple data set.

```{r}
Gender <- rep(c("Male", "Female"), times = 3, each = 4)
Quality <- rep(c("Taste", "Color of package", "Capacity"), times = 1, each = 8)
Accepted <- seq(0, 100, by = 100/23)

DF <- data.frame(Gender, Quality, Accepted)

```

Step 2: Generate boxplot with ggplot

```{r}
library(ggplot2)
ggplot(DF, aes(x = Quality, y = Accepted, fill = Gender)) +
  geom_boxplot() +
  scale_fill_manual(values = c("green", "orange"),
                    labels = expression("Male", "Female")) +
  xlab("") +
  theme(legend.position = "top", legend.title = element_blank()) + 
  guides(fill=guide_legend(reverse=TRUE)) +
  scale_x_discrete(limits=c("Taste", "Color of package", "Capacity"))
```

Step 3: Change the legend order similar to plot order: under themes `fill=guide_legend(reverse=FALSE)`

```{r}
library(ggplot2)
ggplot(DF, aes(x = Quality, y = Accepted, fill = Gender)) +
  geom_boxplot() +
  scale_fill_manual(values = c("green", "orange"),
                    labels = expression("Male", "Female")) +
  xlab("") +
  theme(legend.position = "top", legend.title = element_blank()) + 
  guides(fill=guide_legend(reverse=FALSE)) +
  scale_x_discrete(limits=c("Taste", "Color of package", "Capacity"))
```

step 4: Horizontal bar chart: `coord_flip()`

```{r}
library(ggplot2)
ggplot(DF, aes(x = Quality, y = Accepted, fill = Gender)) +
  geom_boxplot() +
  scale_fill_manual(values = c("green", "orange"),
                    labels = expression("Male", "Female")) +
  xlab("") +
  theme(legend.position = "top", legend.title = element_blank()) + 
  guides(fill=guide_legend(reverse=TRUE)) +
  coord_flip()+
  scale_x_discrete(limits=c("Taste", "Color of package", "Capacity"))
```
