---
title: "Formatting axis labels in ggplot"
author: Thiyanga Talagala
date: '2022-03-27'
categories: [R, tidyverse, ggplot2, scatterplot]
tags: [R, tidyverse, ggplot2]
summary: "With the ggplot2 package, you've got full control over the axes labels in charts. Here we are going to look at some of the most commonly needed formatting options in order to make your graph aesthetically pleasing"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


With the ggplot2 package, you've got full control over the axes labels in charts. Here we are going to look at some of the most commonly needed formatting options in order to make your graph aesthetically pleasing.

## Comma-separated labels

`comma_format` function in `scales` can be used to add comma separated labels. To illustrate this I use African names data set in TidyTuesday.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(scales) 
## read data
african_names <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-16/african_names.csv")
## plot
african_names %>%
  ggplot(aes(year_arrival)) +
  geom_histogram(bins = 20, fill = "darkslateblue", alpha = 0.7, col="white") +
  scale_y_continuous(labels = scales::comma_format()) 
```


## Change label angles, horizontal justification and vertical justification

**Angles only**

The ggplot2 package `element text` can be used to alter the label angles as well as the horizontal and vertical justifications.

```{r, fig.height=5, fig.height=5}
library(palmerpenguins)
penguins %>%
  ggplot(aes(x=species, fill=species)) + geom_bar() +
  theme(axis.text.x=element_text(angle=30))
```

**Angles, horizontal justification and vertical justification**

```{r}
penguins %>%
  ggplot(aes(x=species, fill=species)) + geom_bar() +  scale_fill_manual(values=c("#1b9e77", "#d95f02", "#7570b3")) + 
  theme(axis.text.x=element_text(angle=30, vjust=.8, hjust=0.8))
```





