---
title: "Weighted LOESS (LOcal regrESSion) "
date: "2021-08-07"
slug: blog
categories: [ggplot2]
tags: [R, ggplot2]
---

# `LO`cal regr`ESS`ion (LOESS) or `LO`cally `WE`ighted `S`catter-plot `S`moother (LOWESS)

 LOESS or LOWESS is a nonparametric technique to fit a smooth curve through points in a scatter plot. This approach uses locally estimated linear regression at its core.
 
 The following code illustrates how to include a loess line using the ggplot2 package.
 
 ```{r, comment=NA, warning=FALSE, message=FALSE}
 library(ggplot2)

set.seed(1)
x <- rnorm(60)
y <- c(rnorm(40), 10, rnorm(19))
df <- data.frame(x=x, y=y)

## Without weights
ggplot(data=df, aes(x=x, y=y)) + geom_point() + 
  geom_smooth(method=loess, legend=FALSE)

 ```
 
Furthermore, you can plot loess smoothing weighting the observations as follows:
 
 ```{r, comment=NA, warning=FALSE, message=FALSE}
 
weights <- c(rep(1, 40), 20, rep(1, 19))
df$weights <- weights


ggplot(data=df, aes(x=x, y=y, size=weights, weight=weights)) + geom_point() + 
  geom_smooth(method=loess, legend=TRUE)

 
 ```
 
 
 
 
 