---
title: 'Logistic Regression: Model Building and Interpretation'
author: "Thiyanga Talagala"
date: '2022-04-11'
categories:
- R
- tidyverse
- ggplot2
- scatterplot
tags:
- R
- tidyverse
- ggplot2
summary: 'Logistic regression is a widely used modelling approach, however little
  is known about the modelling processes and interpretation of model outputs. This
  post demonstrates how to build a logistic model with R and how to interpret the
  results. '
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = NA)
```

## Data

In this post I am going to use `penguins` dataset in `palmerpenguins` package in R. First we load the dataset.

```{r, warning=FALSE, message=FALSE}
library(palmerpenguins)
library(skimr)
library(tidyverse)
library(plotly)
library(visdat)
library(patchwork)
data("penguins")

```

Here is an overview of the dataset.

```{r}
skim(penguins)
summary(penguins)
```


Before moving on to the full logistic regression model, it's a good idea to look at the associations between each predictor and gender. Before moving on to the whole model, It's critical to first grasp these linkages. We can use logistic regression to look at all of the predictors simultaneously.



```{r, fig.cap="Distribution of (a) species and (d) island by gender."}
p1_species <- penguins %>% 
  na.omit() %>%
  count(species,sex) %>% 
  group_by(species) %>% 
  mutate(Percentage = 100*n/sum(n)) %>% 
  ggplot(aes(x = species,y = Percentage,fill=sex))+
  geom_bar(stat="identity") + 
  scale_fill_manual(values = c("purple","cyan4")) +
  coord_flip() +
  theme(legend.position = "bottom") +
  ggtitle("a") 

p2_island <- penguins %>% 
  count(island,sex) %>% 
  group_by(island) %>% 
  mutate(Percentage = 100*n/sum(n)) %>% 
  ggplot(aes(x = island,y = Percentage,fill=sex))+
  geom_bar(stat="identity") + 
  scale_fill_manual(values = c("purple","cyan4")) + 
  coord_flip() +
  theme(legend.position = "bottom") +
  ggtitle("b")
(p1_species|p2_island)

```


```{r, fig.cap="Distribution of (a) bill length, (b) bill depth, (c) flipper length and (d) body mass index by gender."}

p3_length <-  ggplot(penguins, aes(x=sex, y=bill_length_mm, fill=sex)) +
  geom_boxplot() + 
  scale_fill_manual(values = c("purple","cyan4")) + 
  theme(legend.position = "bottom") + ggtitle("a")

p4_depth <-  ggplot(penguins, aes(x=sex, y=bill_depth_mm, fill=sex)) +
  geom_boxplot() + 
  scale_fill_manual(values = c("purple","cyan4")) + 
  theme(legend.position = "bottom") + ggtitle("b")

p5_flipper <-  ggplot(penguins, aes(x=sex, y=flipper_length_mm, fill=sex)) +
  geom_boxplot() + 
  scale_fill_manual(values = c("purple","cyan4")) + 
  theme(legend.position = "bottom") + ggtitle("c")

p6_bmi <-  ggplot(penguins, aes(x=sex, y=body_mass_g, fill=sex)) +
  geom_boxplot() + 
  scale_fill_manual(values = c("purple","cyan4")) + 
  theme(legend.position = "bottom") + ggtitle("d")

(p3_length|p4_depth)/(p5_flipper|p6_bmi)
```




## Missing values exploration

```{r}
vis_miss(penguins)
```

## Remove missing values

```{r}
penguins_omit <- penguins %>%
  na.omit()
vis_miss(penguins_omit)
```

## Building logistic regression

Then, using the other variables in the dataset, I'll create a logistic regression model to predict the gender of the palmer penguins. One goal of this work is to identify which variables significantly contribute to determining the gender of penguins. Before beginning to build the model, double-check that all qualitative variables have been saved as factors.


```{r, fig.height=13, fig.width=13}
fit <- glm(sex ~ species + island + bill_length_mm + bill_depth_mm + flipper_length_mm + body_mass_g, data=penguins_omit, family="binomial")
summary(fit)
```

To represent qualitative variables, R will build an indicator (dummy) variable using the lowest coded category as the reference group. For example, 

$$Y=\begin{cases}
0, & \text{if female}\\
1, & \text{if male}
\end{cases}$$

## Interpretation of parameter estimates of logistic regression

The "Estimates" in the above output  corresponds to the **log-odds**. The **odd ratios** can be obtained as follows:

```{r}
exp(coefficients(fit))

```

$$odd = \frac{P(event)}{1-P(event)}$$

The 95\% confidence intervals for log-odds can be obtained as follows

```{r}
confint.default(fit)
```

The 95\% confidence intervals for odds are as follows

```{r}
exp(confint.default(fit))
```


We can write the fitted logistic regression model as follows

$$\hat{Y} = \frac{1}{1+e^{-z}} = \frac{1}{e^{-80.83 -7.40Chinstrap -8.42Gentoo +0.32Dream -0.50Torgersen + 0.61 bill\_length\_mm + 1.64 bill\_depth\_mm + 0.02 flipper\_length\_mm + 0.005 body\_mass\_g}}$$


$$Chinstrap=\begin{cases}
1, & \text{if Chinstrap}\\
0, & \text{otherwise}
\end{cases}$$

$$Gentoo=\begin{cases}
1, & \text{if Gentoo}\\
0, & \text{otherwise}
\end{cases}$$

$$Dream=\begin{cases}
1, & \text{if Dream}\\
0, & \text{otherwise}
\end{cases}$$


$$Torgersen=\begin{cases}
1, & \text{if Torgersen}\\
0, & \text{otherwise}
\end{cases}$$

## Interpretation of parameter estimates

The variables species, bill length (mm), bill depth (mm) and body mass (g) significantly contribute to the model.

With all other factors held constant, this fitted model predicts that the odds of discovering a male penguin on Dream island are 1.38 times higher than on Biscoe island.

For each one-unit increase in bill depth, the probabilities of detecting a male increased 5.18 times (95 percent CI 2.68–10.02). 


> Model Evaluation: Next Post! Model evaluation should be done before interpreting the model. 
