---
title: "Logistic Regression: Model Building and Interpretation"
author: Thiyanga Talagala
date: '2022-04-11'
categories: [R, tidyverse, ggplot2, scatterplot]
tags: [R, tidyverse, ggplot2]
summary: "Logistic regression is a widely used modelling approach, however little is known about the modelling processes and interpretation of model outputs. This post demonstrates how to build a logistic model with R and how to interpret the results. "
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = NA)
```

## Data

In this post I am going to use `penguins` dataset in `palmerpenguins` package in R. First we load the dataset.

```{r, warning=FALSE, message=FALSE}
library(palmerpenguins)
library(skimr)
library(tidyverse)
library(plotly)
library(visdat)
data("penguins")
skim(penguins)
```

Here is an overview of the dataset.

```{r}
summary(penguins)
```

To plot each pair of variables, I used `GGally::ggpairs` function.

```{r}
p1 <- penguins %>%
  select(species, body_mass_g, ends_with("_mm")) %>%
  GGally::ggpairs(aes(color = species)) +
  scale_colour_manual(values = c("darkorange","purple","cyan4")) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"))
ggplotly(p1)
```

## Missing values exploration

```{r}
vis_miss(penguins)
```

## Remove missing values

```{r}
penguins_omit <- penguins %>%
  na.omit()
vis_miss(penguins_omit)
```

## Building logistic regression

Then, using the other variables in the dataset, I'll create a logistic regression model to predict the gender of the palmer penguins. One goal of this work is to identify which variables significantly contribute to determining the gender of penguins. Before beginning to build the model, double-check that all qualitative variables have been saved as factors.


```{r, fig.height=13, fig.width=13}
fit <- glm(sex ~ species + island + bill_length_mm + bill_depth_mm + flipper_length_mm + body_mass_g, data=penguins_omit, family="binomial")
summary(fit)
```


## Interpretation of parameter estimates of logistic regression

The Estimates in the above output  corresponds to the **log-odds**. The **odd ratios** can be obtained as follows:

```{r}
exp(coefficients(fit))

```

The 95\% confidence intervals for log-odds can be obtained as follows

```{r}
confint.default(fit)
```

The 95\% confidence intervals for odds are as follows

```{r}
exp(confint.default(fit))
```


We can write the fitted logistic regression model as follows

$$\hat{Y} = \frac{1}{1+e^{-z}} = \frac{1}{e^{-80.83 -7.40Chinstrap -8.42Gentoo +0.32Dream -0.50Torgersen + 0.61 bill\_length\_mm + 1.64 bill\_depth\_mm + 0.02 flipper\_length\_mm + 0.005 body\_mass\_g}}$$

## Interpretation of parameter estimates

The variables species, bill length (mm), bill depth (mm) and body mass (g) significantly contribute to the model.

Holding all other variables constant, this fitted model predicts that the odds of detecting a female are 1.38 times higher for penguins on Dream Island than on Biscoe.


The odds of detecting a female increased 1.005 times (95%CI 1.003–1.007) for each one-unit
increase in the body mass (g).

> Model Evaluation: Next Post! Model evaluation should be done before interpreting the model. 
