---
title: "Logistic Regression: Model Building and Interpretation"
author: Thiyanga Talagala
date: '2022-04-11'
categories: [R, tidyverse, ggplot2, scatterplot]
tags: [R, tidyverse, ggplot2]
summary: "Logistic regression is a widely used modelling approach, however little is known about the modelling processes and interpretation of model outputs. This post demonstrates how to build a logistic model with R and how to interpret the results. "
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = NA)
```

## Data

In this post I am going to use `penguins` dataset in `palmerpenguins` package in R. First we load the dataset.

```{r, warning=FALSE, message=FALSE}
library(palmerpenguins)
library(skimr)
library(tidyverse)
library(plotly)
library(visdat)
data("penguins")
skim(penguins)
```

Here is an overview of the dataset.

```{r}
summary(penguins)
```

To plot each pair of variables, I used `GGally::ggpairs` function.

```{r}
p1 <- penguins %>%
  select(species, body_mass_g, ends_with("_mm")) %>%
  GGally::ggpairs(aes(color = species)) +
  scale_colour_manual(values = c("darkorange","purple","cyan4")) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"))
ggplotly(p1)
```

## Missing values exploration

```{r}
penguins_omit <- penguins %>%
  na.omit()
vis_miss(penguins_omit)
```

## Remove missing values

## Building logistic regression

Then, using the other variables in the dataset, I'll create a logistic regression model to predict the gender of the palmer penguins. Before beginning to build the model, double-check that all qualitative variables have been saved as factors.


```{r}
fit <- glm(sex ~ species + island + bill_length_mm + bill_depth_mm + flipper_length_mm + body_mass_g, data=penguins_omit, family="binomial")
summary(fit)
```


## Interpretation of parameter estimates of logistic regression

The Estimates in the above output  corresponds to the **log-odds**. The **odd ratios** can be obtained as follows:

```{r}
exp(coefficients(fit))

```

The 95\% confidence intervals for log-odds can be obtained as follows

```{r}
confint.default(fit)
```

The 95\% confidence intervals for odds are as follows

```{r}
exp(confint.default(fit))
```


We can write the estimated logistic regression model as follows

$$\phi(z) = \frac{1}{1+e^{-z}} = \frac{1}{e^{-80.83 -7.40Chinstrap -8.42Gentoo +0.32Dream -0.50Torgersen + 0.61 bill\_length\_mm + 1.64 bill\_depth\_mm + 0.02 flipper\_length\_mm + 0.005 body\_mass\_g}}$$