---
title: "How to install and load multiple packages at once?"
author: Thiyanga Talagala
date: '2019-03-22'
categories: [R, tidyverse, gganimate]
tags: [R, tidyverse, gganimate]
summary: "Install multiple R packages at once/ check and install missing packages."
---

<style type="text/css">
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
</style>

## install packages

To install multiple packages with a single call to `install.packages`, pass the names of the packages as a character vector to the install.packages() function.

```{r, warning=FALSE, message=FALSE, eval=FALSE}
install.packages(c("gganimate", "tidyverse", "gapminder"))
```


## load packages

Once you have the packages installed, you can make their contents available to use in your current R session by running,

```{r,warning=FALSE, message=FALSE}
lapply(c("gganimate", "tidyverse", "gapminder"), require, character.only = TRUE)
```


## check and install missing packages


```{r,warning=FALSE, message=FALSE, eval=FALSE}
list.of.packages <- c("gganimate", "tidyverse", "gapminder")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
```


## For other alternatives see

1. [littler](http://dirk.eddelbuettel.com/code/littler.html)

```r
library(littler)
install.r EIAdata gdata ggmap ggplot2 
```


2. [ipak](https://gist.github.com/stevenworthington/3178163)

```{r,warning=FALSE, message=FALSE}
ipak <- function(pkg){
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}

# usage
packages <- c("gganimate", "tidyverse", "gapminder")
ipak(packages)

```

3. [easypackages](https://cran.r-project.org/web/packages/easypackages/README.html)