---
title: "Some useful functions to Wrangle with Time Series Data"
author: Thiyanga Talagala
date: '2023-03-14'
categories: [R, tidyverse, tsibble, time series]
tags: [R, tidyverse, tsibble, time series]
summary: "Time series data wrangling"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=TRUE, echo = TRUE, comment = NA)
```

# Creating data and time objects.

The `lubridate`  package in R provides a set of functions for working with dates and times. 

## year-month-date `ymd` function

`ymd` function converts a character vector of dates in the format "year-month-day" to a Date object. The name of the function stands for "year-month-day", which is the order of the components in the date format that the function expects. 

```{r}
library(lubridate)
dates.ex1 <- c("2022-03-15", "2022-03-16", "2022-03-17")
class(dates.ex1)
dates1 <- ymd(dates.ex1)
dates1
class(dates1)

```

The ymd() function also supports variations of the "year-month-day" format, such as "year/month/day" or "year.month.day". 

```{r}
dates.ex2 <- c("2022/03/15", "2022/03/16", "2022/03/17")
dates2 <- ymd(dates.ex2)
dates2
class(dates2)
```



```{r}
dates.ex3 <- c("2022/03/15", "2022/03/16", "2022/03/17")
dates2 <- ymd(dates.ex2)
dates2
class(dates2)
```

## month-date-year `mdy` function

```{r}
mdy(c("03-15-2023", "03-16-2023"))

```

## date-month-year `dmy` function

```{r}
dmy(c("15-03-2023", "16-03-2023"))
```

## dates with strings

```{r}
ydm("2023, March 15th")
mdy("March 15th, 2023")
mdy("Mar 15th, 2023")
mdy("Wednesday, March 15th, 2023")
mdy("15th, March, 2023")
```

## hour-minutes-second `hms` working with dates and times

```{r}
hms("10:30:00")
```

## `ymd_hm`, `ymd_hms` and `mdy_hms`

```{r}
ymd_hm("2023-03-15 10:30")
ymd_hms("2023-03-15 10:30:00")
mdy_hms("03-15-2023 10:30:00")
```


##  `parse_date_time`

If you have a date format that doesn't match any of the built-in formats, you can use the parse_date_time() function in lubridate to specify a custom format.

```{r}
dates.ex4 <- c("2022/03/15 10:30:00", "2022/03/16 11:00:00", "2022/03/17 12:15:00")

# Define the custom format using the format string
date_format <- "%Y/%m/%d %H:%M:%S"

# Convert the character vector to POSIXct objects using parse_date_time()
parse_date_time(dates.ex4, date_format)


```



# Creating a sequence of time

```{r}
y1 <- seq(as.Date("2012-01-01"),as.Date("2012-12-01"), by = "1 month")
y1
class(y1)
```

```{r, echo=TRUE, eval=TRUE}
seq(as.Date("2023-03-01"),as.Date("2023-04-01"), by = "1 week")
```



# Extract various time components

Let's see how to extract various time components from the elements of the following Date object.

```{r}
y2023 <- seq(as.Date("2023-01-01"),as.Date("2023-12-31"), by = "1 month")
```

`year()`: Extracts the year component 

```{r}
year(y2023)
```

`quarter()`: Extracts the quarter 

```{r}
quarter(y2023)
```

`month()`: Extracts the month component 

```{r}
month(y2023)
```

`week()`: Extracts the week component 

```{r}
week(y2023)
```


day(): Extracts the day component 

```{r}
day(y2023)
```


To illustrate the remaining functions I use the following example

```{r}
mydate <- ymd_hms("2023-03-15 10:30:01")
```

hour(): Extracts the hour component

```{r}
hour(mydate)
```


minute(): Extracts the minute component 

```{r}
minute(mydate)
```

second(): Extracts the second component 

```{r}
second(mydate)
```

