---
title: "Data Wrangling"
author: "Thiyanga Talagala"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, comment = NA)
```


Load the `gapminder` dataset and the `tidyverse` and `magrittr` packages.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(gapminder)
library(tidyverse)
library(magrittr)
```


# Exercises

1. Filter all rows for "Sweden".

```{r}
filter(gapminder, country=="Sweden")

```

2. Filter all rows where `lifeExp` is less than or equal to 30.

```{r}
gapminder %>% filter(lifeExp <= 50)
```

3. Filter all rows that have a missing value for `year`.

```{r}
filter(gapminder, is.na(year))
```

4. Filter all countries that had population over 100000 in 1960 or earlier.

```{r}
filter(gapminder, pop>100000 & year <=1960)

```

5. Count the number of countries with life expectancy greater than 30 in 1952.

```{r}
df <- gapminder %>%
  filter(year==1952 & lifeExp < 30) 
df
dim(df)
```

6. Calculate the mean life expectancy for each year and continent.

```{r}
gapminder %>%
  group_by(continent, year) %>%
  summarise(mean.lifeExp = mean(lifeExp))

```

7. Get the maximum and minimum of GDP per capita for all continents in a "wide" format.

```{r}
gapminder %>%
  group_by(continent) %>%
  summarize(maxGdpPercap=max(gdpPercap),
           minGdpPercap=min(gdpPercap))
```

8. Get the maximum and minimum of GDP per capita for all continents in a "long" format.

```{r}
gapminder %>%
  group_by(continent) %>%
  summarize(maxGdpPercap=max(gdpPercap),
           minGdpPercap=min(gdpPercap)) %>%
  pivot_longer(2:3, "summary", "value")
```

9. What was the population of the United States in 1952 and 2007.

```{r}
gapminder %>%
  filter(country=="United States", year %in% c(1952, 2007)) 

```

10. Subset the gapminder data to extract rows where `lifeExp` is greater than or equal 80. Retain only the columns `country`, `year`, and `lifeExp`. Sort the results from largest to smallest based on `lifeExp`.

```{r}
gapminder %>%
filter(lifeExp >= 80) %>%
select(country, year, lifeExp) %>%
arrange(desc(lifeExp))
```

11. Calculate the total GDP in billions of dollars, extract the results for the year 2002, and sort the rows so that the total GDP is in decreasing order.

Help: `gpd = gdpPercap * pop`

```{r}
gapminder %>%
mutate(gdp = gdpPercap * pop) %>%
  filter(year==2002) %>%
  arrange(desc(gdp))
```

12. Calculate the average life expectancy by continent in 2002.

```{r}
gapminder %>%
  filter(year==2002) %>%
group_by(continent) %>%
summarize(mean_lifeExp=mean(lifeExp))
```

13. Which countries and which years had the worst five GDP per capita measurements?

```{r}
gapminder %>%
  arrange(desc(gdpPercap)) %>%
  tail(5)

```

14. What was the mean life expectancy across all countries for each year in the dataset?

```{r}
gapminder %>%
  group_by(year) %>%
  summarize(mean(lifeExp))
```

15. Which five Asian countries had the highest life expectancy in 2007?

```{r}
gapminder %>%
  filter(continent=="Asia") %>%
  arrange(desc(lifeExp)) %>%
  head(5)
```

16. Calculate the total number of observations for each country in Europe. Help: use `n()` function.

```{r}
gapminder %>%
filter(continent == "Europe") %>%
group_by(country) %>%
summarize(n = n())
```

17. How many observations do we have per continent?

```{r}
gapminder %>%
group_by(continent) %>%
summarize(n = n())
```

18. Compute the average life expectancy by
continent.

```{r}
gapminder %>%
group_by(continent) %>%
summarize(avg_lifeExp = mean(lifeExp))

```

19. Rank countries according to their life expectancy and store it in a new column called rank. Rearrange the rows according to the ascending order of ranks (1, 2, 3...).


```{r}
gapminder %>% 
  filter(year == 2007) %>% 
  select(country, lifeExp) %>% 
  mutate(rank = min_rank(desc(lifeExp))) %>%
  arrange(rank)
```

20. Calculate the mean and the standard error of the life expectancy for Belgium, Netherlands and France.

```{r}
gapminder %>% 
  filter(country %in% c("Belgium", "Netherlands", "France")) %>%
  group_by(country) %>%
  summarize(mean = mean(lifeExp), se = sd(lifeExp)/sqrt(n()))
```

21. Categorize countries as "low" (`lifeExp` < 50) and "high" (`lifeExp` > 50) and store the values in a new column named "category".
 
```{r}
 gapminder %>%
    mutate(category = ifelse(lifeExp > 50, "high", "low")) 
```